---
title: "EDLD652_final"
author: "Hyeonjin Cha"
date: "2/24/2021"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(rio)
library(here)
library(janitor)
library(usmap)
library(ggpubr)
library(reactable)
library(magrittr)
```


```{r load data}
covid <- import(here("data", "covid-data-01-30-2021.csv")) %>% 
  clean_names() %>% 
  as_tibble() #load covid data

soccap <- import(here("data", "social-capital-project-state-index.xlsx")) %>% 
  clean_names() %>%
  as_tibble() #load state level social capital data 

soccap_covid <- soccap %>% 
  left_join(covid, by = c("state" = "province_state"))

soccap_county <- import(here("data", "social-capital-project-county-index.xlsx")) %>% 
  clean_names() %>% 
  as_tibble() %>% 
  rename(fips = fips_code) #load county level social capital data 

```

Data Source {.sidebar}
==========
**Data Source**

These figures use data from two different sources:

1) State-Level Social Capital Index: Social Capital Project by the United States Congress Joint Economic Committee (https://www.jec.senate.gov/public/index.cfm/republicans/2018/4/the-geography-of-social-capital-in-america)

2) State-Level Covid-19 Statistics: COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University (https://github.com/CSSEGISandData/COVID-19)

**Visualization #1**

Covid-19 testing, cases, and deaths by each state.

**Visualization #2**

Mapping state-level social Capital index.

**Visualization #3**

Regressions examining the relationship between social capital index and Covid-19 metrics.



Raw Data {data-icon="fa-table"}
=====================================     
### Raw Data
```{r}
soccap_covid_clean <- soccap_covid %>% 
  select(state, fips_code, 
         state_level_index, 
         family_unity, family_interaction, social_support, community_health, institutional_health, collective_efficacy, philanthropic_health,
         total_test_results,
         confirmed,
         deaths
         )
reactable(soccap_covid_clean)
```

Visualization #1: Covid-19 {data-icon="fa-signal"}
=====================================     

Column 1 {data-width=400}
-----------------------------------------------------------------------
### Candidate A

```{r}
#I'm a fan of Candidate A, since you can see the different configuration of the states and the difference in scales makes it almost impossible to see deaths in Candidate B. Reisizing the graph to make it taller will help viewers read the state names bette, and angling the x axis and reducing the text elements (I believe R will automatically do this if you change the variable to a factor) will also improve readability. I like the use of colors! 

#Plot part 1. Covid cases, deaths, and testing by state
data_plota <- soccap_covid_clean %>%
  select(state, confirmed, deaths, total_test_results)

state_covid_test <- data_plota %>%
  mutate(state = fct_reorder(state, total_test_results))  %>% 
  ggplot(aes(state, total_test_results/1000)) +
  geom_col(fill = "deepskyblue3", alpha = 0.7) + 
  coord_flip() +
  labs(x = "State", subtitle = "Tests", y="") + 
  theme_minimal()

state_covid_case <- data_plota %>%
  mutate(state = fct_reorder(state, confirmed))  %>% 
  ggplot(aes(state, confirmed/1000)) +
  geom_col(fill = "darkorange", alpha = 0.7) + 
  coord_flip() +
  labs(x = " ", subtitle = "Cases", y = "Counts by 1000") + 
  theme_minimal()

state_covid_death <- data_plota %>%
  mutate(state = fct_reorder(state, deaths))  %>% 
  ggplot(aes(state, deaths/1000)) +
  geom_col(fill = "brown1", alpha = 0.7) + 
  coord_flip() + 
  labs(x = " ", subtitle = "Deaths", y = "") + 
  theme_minimal()

covid_plots <- ggarrange(state_covid_test,
                         state_covid_case,
                         state_covid_death,
                         ncol = 3,
                         nrow = 1)
covid_plots

```

Column 2 {data-width=400}
-----------------------------------------------------------------------
### Candidate B
```{r}
#Plot part 1. Covid cases, deaths, and testing by state
state_covid_altogether <- data_plota %>%
  mutate(state = fct_reorder(state, total_test_results))  %>% 
  pivot_longer(!state, names_to = "type", values_to = "count") %>%
  ggplot(aes(state, count/1000, fill = type)) +
  geom_col(position = "dodge2") + 
  scale_fill_brewer(palette = "Dark2") + 
  coord_flip() + 
  theme_minimal()
state_covid_altogether

```

Visualization #2: Social Capital Index {data-icon="fa-signal"}
===================================== 

Column 1 {data-width=800}
-----------------------------------------------------------------------
### State Level Index
    
```{r}
#I think it might look nicer to display these in a row rather than a column. Making the graphs interactive, so that the names only show up if you hover over the state or county would fix the problem with overlapping labels. 

plot_usmap(data = soccap_covid_clean, regions = "states", values = "state_level_index", labels = TRUE) +
  scale_fill_continuous(low = "white", 
                        high = "steelblue1") + 
  theme(panel.background = element_rect(color = "black", fill = "white")) + 
  theme(legend.position = "none") + 
  labs(title = "State-Level Overall Social Capital Index",
       subtitle = "Thicker color indicates higher levels of social capital index")

```
    
### County Level Index in Oregon

```{r}
plot_usmap(data = soccap_county, include = c("OR"), 
           regions = "counties", values = "county_level_index", labels = TRUE) +
  scale_fill_continuous(low = "white", 
                        high = "steelblue1") + 
  theme(panel.background = element_rect(color = "black", fill = "white")) + 
  theme(legend.position = "none") + 
  labs(title = "County-Level Overall Social Capital Index in Oregon",
       subtitle = "Thicker color indicates higher levels of social capital index")

```

Visualization #3: Regressions {data-icon="fa-signal"}
===================================== 
Column 1 {data-width=400}
-----------------------------------------------------------------------

```{r}
#Love the regression print out at the bottom! Here, it could be helpful to provide more detail about what the State Level Social Capital Index is. 

soccap_covid_clean %<>%
  mutate(positivity = confirmed / total_test_results, fatality = deaths / confirmed) 

soccap_covid_clean %>% 
  ggplot(aes(state_level_index, positivity)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "black", se = FALSE) + 
  labs(title = "Relationship between Social Capital and Covid-19 Positivity",
       x = "State Level Social Capital Index",
       y = "Posivitity Rate (%)") + 
  theme(panel.background = element_rect(color = "black", fill = "white")) 
```

```{r}
model1a <- lm(positivity ~ state_level_index, data = soccap_covid_clean)
summary(model1a)
```

Column 2 {data-width=400}
-----------------------------------------------------------------------
```{r}
model2a <- lm(fatality ~ state_level_index, data = soccap_covid_clean)
model2b <- lm(fatality ~ family_unity + family_interaction + social_support + community_health + institutional_health + collective_efficacy + philanthropic_health, data = soccap_covid_clean)

soccap_covid_clean %>% 
  ggplot(aes(state_level_index, fatality)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "black", se = FALSE) + 
  labs(title = "Relationship between Social Capital and Covid-19 Fatality",
       x = "State Level Social Capital Index",
       y = "Fatality Rate (%)") + 
  theme(panel.background = element_rect(color = "black", fill = "white")) 
```

```{r}
model2a <- lm(fatality ~ state_level_index, data = soccap_covid_clean)
summary(model2a)
```
