---
title: "EDLD652_final"
author: "Hyeonjin Cha"
date: "2/24/2021"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(rio)
library(here)
library(janitor)
library(usmap)
library(ggpubr)
library(reactable)
```


```{r load data}
covid <- import(here("data", "covid-data-01-30-2021.csv")) %>% 
  clean_names() %>% 
  as_tibble() #load covid data

soccap <- import(here("data", "social-capital-project-state-index.xlsx")) %>% 
  clean_names() #load social capital data for 1997

soccap_covid <- soccap %>% 
  left_join(covid, by = c("state" = "province_state"))

```

Data Source {.sidebar}
==========
**Data Source**

These figures use data from two different sources:

1) State-Level Social Capital Index: Social Capital Project by the United States Congress Joint Economic Committee (https://www.jec.senate.gov/public/index.cfm/republicans/2018/4/the-geography-of-social-capital-in-america)

2) State-Level Covid-19 Statistics: COVID-19 Data Repository by the Center for Systems Science and Engineering (CSSE) at Johns Hopkins University (https://github.com/CSSEGISandData/COVID-19)

**Visualization #1**

jj


**Visualization #2**

jj


**Visualization #3**

jj



Raw Data {data-icon="fa-table"}
=====================================     
### Raw Data
```{r}
soccap_covid_clean <- soccap_covid %>% 
  select(state, fips_code, 
         state_level_index, 
         family_unity, family_interaction, social_support, community_health, institutional_health, collective_efficacy, philanthropic_health,
         total_test_results,
         confirmed,
         deaths
         )
reactable(soccap_covid_clean)
```

Visualization #1: Covid-19 {data-icon="fa-signal"}
=====================================     

Column 1 {data-width=700}
-----------------------------------------------------------------------
### Candidate A

```{r}
#Plot part 1. Covid cases, deaths, and testing by state
data_plota <- soccap_covid_clean %>%
  select(state, confirmed, deaths, total_test_results)

state_covid_test <- data_plota %>%
  mutate(state = fct_reorder(state, total_test_results))  %>% 
  ggplot(aes(state, total_test_results/1000)) +
  geom_col(fill = "deepskyblue3", alpha = 0.7) + 
  coord_flip() +
  labs(x = "State", subtitle = "Tests", y="") + 
  theme_minimal()

state_covid_case <- data_plota %>%
  mutate(state = fct_reorder(state, confirmed))  %>% 
  ggplot(aes(state, confirmed/1000)) +
  geom_col(fill = "darkorange", alpha = 0.7) + 
  coord_flip() +
  labs(x = " ", subtitle = "Cases", y = "Counts by 1000") + 
  theme_minimal()

state_covid_death <- data_plota %>%
  mutate(state = fct_reorder(state, deaths))  %>% 
  ggplot(aes(state, deaths/1000)) +
  geom_col(fill = "brown1", alpha = 0.7) + 
  coord_flip() + 
  labs(x = " ", subtitle = "Deaths", y = "") + 
  theme_minimal()

covid_plots <- ggarrange(state_covid_test,
                         state_covid_case,
                         state_covid_death,
                         ncol = 3,
                         nrow = 1)
covid_plots

```

Column 2 {data-width=700}
-----------------------------------------------------------------------
### Candidate B
```{r}
#Plot part 1. Covid cases, deaths, and testing by state
state_covid_altogether <- data_plota %>%
  mutate(state = fct_reorder(state, total_test_results))  %>% 
  pivot_longer(!state, names_to = "type", values_to = "count") %>%
  ggplot(aes(state, count/1000, fill = type)) +
  geom_col(position = "dodge2") + 
  scale_fill_brewer(palette = "Dark2") + 
  coord_flip() + 
  theme_minimal()
state_covid_altogether

```

Visualization #2: Social Capital Index {data-icon="fa-signal"}
===================================== 

Column 1 {data-width=700}
-----------------------------------------------------------------------
### State Level Index
    
```{r}
plot_usmap(data = soccap_covid, regions = "states", values = "state_level_index", labels = TRUE) +
  scale_fill_continuous(low = "white", 
                        high = "steelblue1") + 
  theme(panel.background = element_rect(color = "black", fill = "white")) + 
  theme(legend.position = "none") + 
  labs(title = "State-Level Overall Social Capital Index",
       subtitle = "Thicker color indicates higher levels of social capital index")


```
    
### Chart 2

```{r}
```

-----------------------------------------------------------------------

### Table

```{r}
#I tried faceting cases, deaths and testing but the number for testing is too big compared to two other metricss


```

### Chart C

```{r}

```

